Conversational expectations account for apparent limits on theory of mind use

### 2016-01-30 writing/cogsci-submission/cogsci_keysar.tex
Shrunk abstract, formatting

\begin{abstract}
Theory of mind is a powerful cognitive ability: by the age of six, people are capable of accurately reasoning about the beliefs and desires underlying other people's behavior. An influential series of language understanding experiments by Keysar and colleagues, however, showed that adults systematically failed to take the speaker's beliefs into account. This has been widely interpreted as evidence for limitations on theory of mind use. In this paper we argue that these apparent failures are in fact successes. Pragmatic language understanding requires sophisticated social reasoning about the speaker's goals and context. Through a minimal pair of replications comparing scripted vs. unscripted speakers, we show that critical utterances used by Keysar and colleagues are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation. Listeners who pragmatically expect more informative utterances are thus tricked by confederates into making systematic patterns of errors. When we allow participants to naturally interact, we find that their expectations are justified and errors are reduced. This ironically shows that apparent failures of theory of mind are in fact attributable to sophisticated expectations about speaker behavior---that is, to theory of mind.

@@

\begin{figure*}[bt!]
\begin{center}
\includegraphics[scale = .4]{images/overall_view.png}
\end{center}
\vspace{-.25cm}
\caption{Interface used in the reported experiments. Objects behind the black squares were hidden from the director.}
\label{fig:interface}
\end{figure*}

@@

The argument of Keysar and colleagues is based on an elegant experimental paradigm, where participants played a simple communication game with a confederate. The two players were placed on opposite sides of a $4 \times 4$ grid containing a set of everyday objects (see Fig. \ref{fig:interface}). The confederate played the role of `director,' giving instructions about how to move objects around a grid, and the participant played the role of `matcher,' attempting to follow these instructions. For example, the objects in one trial included a cassette tape. The director gave an instruction like `move the tape up one square,' referring to the cassette.
Critically, some objects were occluded such that only the matcher could see them, creating an asymmetry in the players' knowledge. To perform accurately on critical trials, the matcher would need to apply theory of mind to reason about the director's beliefs. For example, imagine a roll of tape were placed in an occluded slot: if a participant failed to account for the director's (partial) knowledge, she might interpret `tape' to mean the occluded roll of tape (which the director couldn't possibly know about). 
Indeed, \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} found that participants attempted to move the hidden item in 30\% of cases: 71\% of participants attempted to move this hidden item at least once (out of four critical cases) in the experimental condition, compared to 0\% in a control condition where there was no ambiguity over the referent. 
Additionally, eye-tracking data showed that participants considered the hidden item more often and for longer in the experimental condition than the control condition.

While these results are compelling, the paradigm has been criticized from a couple different angles. 
First, \citeA{HellerGrodnerTanenhaus08_Perspective} have pointed out that in many cases, the hidden object was a better fit for the referring expression than the one in common ground (e.g. the hidden roll of tape vs. the cassette tape for ``the tape''), making the hidden object \emph{a priori} more likely to be the referent; it would generally be fairer to compare two objects that fit the referring expression equally well. 
We validate this argument by empirically measuring relative fit of the expressions to the target and distractor items.
% Although relative fit was not empirically evaluated prior to our study,
Second, the viewpoint asymmetry paradigm is somewhat unnatural: common ground is typically built incrementally over the course of an interaction rather than presented all at once, and it is rare for a shared display to differ in perceptual accessibility \cite{HannaTanenhausTrueswell03_CommonGroundPerspective}. 
%Third, none of the experiments reported by Keysar and colleagues included a key comparison condition where the critical item (e.g. the roll of tape) was \emph{also} in common ground \cite{BrownSchmidtHanna11_IncrementalPerspectiveTaking}---perhaps people would refer to the critical item less in the privileged condition than in the full common ground condition, which would put a more positive spin on the results. %Note that none of these criticisms invalidate the paradoxical result that listeners move an occluded object, they just suggest an alternative explanation. Reference disambiguation is probabilistic: common ground and goodness of fit are weighed against one another.

In this paper, we offer an additional factor that helps account for Keysar's results. 
Theory of mind as applied within language understanding depends on an accurate model of what a speaker would say in different situations; given an utterance, a listener can then reason backward to the most plausible situation \cite{Grice75_LogicConversation,Clark96_UsingLanguage,FrankGoodman12_PragmaticReasoningLanguageGames, GoodmanStuhlmuller13_KnowledgeImplicature}.
This suggests that we consider whether the utterances produced by the confederate in Keysar's critical conditions were actually what a speaker in that context would be expected to say. 
If not, then perhaps listeners are making choices that are in fact consistent with a correct pragmatic interpretation of the confederate's (uncooperative) utterance.
More precisely, when both players know that objects are occluded in the display, the speaker may tend to add additional precision to references in order to avoid confusion. If the listener \emph{expects} the speaker to do this, they will pragmatically pick the \emph{a priori} more likely referent of the referring expression, which in critical trials will be the occluded object. In other words, it is precisely \emph{because} the listener takes the speaker's mental state into consideration that they are tricked by an uncooperative confederate into choosing the wrong item. 
\begin{table*}
\begin{center}
\begin{tabular}{ p{2cm} | r | r |  r || r | r | r || r | r | r}
& \multicolumn{3}{c||}{\% attempted at least once} & \multicolumn{3}{c||}{\% attempted at least twice} & \multicolumn{3}{c}{\% of total cases}\\
\hline
& Orig. & Expt. 1 & Expt. 2 & Orig. & Expt. 1 & Expt. 2 & Orig. & Expt. 1 & Expt. 2  \\
Experimental & 71 & 93 & 61 & 46 & 57 & 32 & 30 & 43 & 24\\
Baseline        & 0   & 7   & 0   & 0   & 0   & 0   & 0   & 2   & 0\\
\end{tabular}
\caption{Side-by-side comparison of error rates in \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} and our two replications. The first section shows the percentage of \emph{participants} attempting to move the occluded distractor at least one of the four possible cases. The second section shows the percentage attempting to move it at least twice. The third section shows the percentage of \emph{all experimental trials} that the participant actually tried to move the occluded object. }
\label{table:mainResults}
\end{center}
\end{table*}

We began by replicating \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in a multi-player web experiment. We recruited participants to be both director and matcher (instead of using a confederate), but instructions for critical items, as well as a random subset of filler items, remained scripted as in the original study. We replicated the original finding, but noted a tendency of directors to be overinformative in unscripted trials. We then ran the same experiment without scripting the directions for critical trials, observing unconstrained director behavior. We observe much greater precision in unconstrained director utterances, which match targets much better than distractors, and better performance of the matchers. This minimal pair of experiments demonstrates that listener mistakes are at least partially due to the pragmatics of the task, ironically showing that apparent failures of theory of mind are in fact attributable to sophisticated expectations about speaker behavior---that is, to theory of mind.

@@

This paradigm differs from those used by \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in three primary ways. First, participants were not seated across from each other at a table: they each saw a views of the 4 x 4 grid on screen and communicated via a text box. Second, we did not use a trained confederate. We randomly assigned one of the players to the role of the instruction giver, and maintained the original wording by scripting a subset of their instructions. Finally, the hidden object was not placed in a bag, in which respect our design more closely resembles \citeA{KeysarBarr___Brauner00_TakingPerspective}.

%The experiment proceeded through all eight items, with the director providing instructions and the participant moving the objects. Except for instructions with scripted messages, the participants were free to use the chat box for questions, small talk, or whatever else.
\begin{table*}
\begin{center}
\begin{tabular}{ m{.75cm} | m{1.5cm} | m{1.5cm} | m{1.5cm} |  m{1.25cm} | m{1.6cm} | m{1.4cm} | m{1.3cm} | m{1.3cm} | m{1.5cm} |}
& & Item 1 & Item 2 & Item 3 & Item 4 & Item 5 & Item 6 & Item 7 & Item 8\\\hline
& instruction & ``glasses'' & ``bottom block'' & ``tape'' & ``large measuring cup'' & ``brush'' & ``eraser'' & ``small candle'' & ``mouse'' \\\hline
& target & sunglasses & block (3rd row) & cassette & medium cup & round hairbrush & board eraser & medium candle & computer mouse \\\hline
& hidden distractor & glasses case & block (4th row) & scotch-tape & large cup & flat \,\,\,hairbrush & pencil eraser & small candle & toy mouse \\
 \hline\hline
\multirow{2}{.75cm}{Expt.~1}  & \# incorrect & 0 & 5 & 1 & 3 & 2 &6 & 6 & 1 \\ 
                                               & \# correct & 6 & 3 & 1 & 8 & 2 & 2 & 4 & 6 \\
 \hline\hline
\multirow{2}{.75cm}{Expt.~2} & \# incorrect & 0 & 1 & 1 & 3 & 9 & 7 & 1 & 5 \\
                                              & \# correct   & 12&14&11&13&7 &8   &12& 8 \\

\end{tabular}
\caption{Item-wise error rates for critical trials}
\label{table:ItemWise}
\end{center}
\end{table*}

@@

As a proxy for the eye-tracking analyses reported by \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, we conducted a mouse-tracking analysis. We define the decision window as the span of time between the point when the matcher received their instruction message and when they started moving an object. If it took them multiple attempts to move the correct object, we restricted our analysis to the first attempt. We took the set of all mouse position samples in the decision window (collected approximately every 15ms), and measured what proportion were located inside the cell containing the target vs. the cell containing the distractor.  We had to exclude an additional 3 participants for this analysis, because the timestamps for director and matcher did not align and we could not establish the decision window properly.
We found that people tended to spend less total time hovering over the target cell on critical experimental trials than on baseline trials, $b = -.11, t(20) = -1.26, p = 0.224$, although perhaps due to an underpowered sample and imprecise measurements\footnote{This analysis includes participants who actually made errors, since the data is too sparse to exclude them.}, this result was not significant (see Fig. \ref{fig:hover}). %accounts for a significant proportion of variance $F(3, 52) = 10.7, p < 0.001$

By running this replication as a multi-player web experiment we have available an additional source of data beyond the original experiments: half of the instructions were unscripted, providing observations of natural production of referential descriptions for filler items. 
Informally, we noted a tendency toward, possibly unnecessary, precision in descriptions. Instead of ``move the stuffed animal down'', participants say ``move the stuffed panda bear down.'' Or, instead of saying ``move the plane to the right'' when there is only one plane, participants say ``move the red airplane to the right.'' 
Perhaps directors were taking the time to make more precise descriptions because they believed it was contextually relevant: both parties know that there are hidden objects in the environment increasing the chance of miscommunication from imprecise descriptions. 
If the matcher expected the director to be precise, then they would be justified in picking the first or best object that meets the description (rather than worrying excessively about occluded cells). 
That is, the scripted instructions used by the director for critical trials may have been \emph{uncooperative} for this situation, and thus led matchers astray. 
We tested this prediction in Expt.~2, where we removed the scripted instructions and allowed speakers to refer to items however they wished.
By the reasoning above we expected to see more precise descriptions by unscripted directors and fewer errors by matchers in the critical trials.

@@

 \begin{figure}[b!]
\begin{center}
\includegraphics[scale=.5]{images/mousetracking.pdf}
\caption{Mean percentage of time spent hovering over target cell in the experimental and baseline conditions for Expt.~1 (left) and Expt.~2 (right). Error bars are 95\% confidence intervals.}
\label{fig:hover}
\end{center}
\end{figure}

@@

We find that patterns of errors in Expt.~2 diverge significantly from a uniform distribution across items, $\chi^2(7) = 24.8, p < 0.001$. Looking more closely at these patterns, we see that the only items where errors are consistently made are those where the more precise utterances used remain ambiguous. For example, in item 5, both the target and distractor are hair brushes: one is round and one is flat. ``Hair brush'' is already a label at the sub-class level. If we remove the two most difficult items (the ``hair brushes'' and the ``erasers''), the total percentage of errors on experimental trials drops from 24\% to 10\% and the percentage of participants making at least one error drops from 61\% to 32\%.

When we conducted a mouse-tracking analysis identical to the one reported for Expt.~1, we again found a significant decrease in total hover time over the target in experimental vs. baseline trials $b = 0.07, t(54) = -2.1, p = 0.045$, indicating that the presence of a hidden distractor still interfered with participants directly choosing the target. While the effect appears to be attenuated compared to Expt.~1 (see Figure \ref{fig:hover}), there is no significant interaction of experiment and condition on hover time $b = 0.04, t(74) = 0.45, p = 0.657$. Again, given the low sample size in Expt.~1, the noise from error, and the noise from counting samples instead of times, this result is limited.
\begin{figure}[t!]
\begin{center}
\includegraphics[scale=.5]{images/itemwisefitness.pdf}
\caption{Item-wise error rates compared with label fitness ratios, for Expt.~1. Fitness ratios are scaled to the unit interval in order to visually compare to error rates. Error bars are bootstrapped 95\% confidence intervals for fitness and 95\% highest density intervals for error rates.}
\label{fig:itemLevel}
\end{center}
\end{figure}

Next, we test whether these improvements in performance are in fact due to more informative speaker behavior. We recruited twenty judges on Amazon Mechanical Turk, who provided ratings for how well the 71 unique labels used by speakers across both experiments (including scripted labels) fit the target and hidden distractor objects. Their responses were given on a slider with endpoints labeled ``not at all'' and ``perfectly.''  Inter-rater reliability was relatively high, with intra-class correlation coefficient of $0.6$ (95\% CI = $[0.54, 0.66]$). In a mixed model including random intercepts for raters and items, we found a significant crossover interaction of experiment and referent on mean fitness rating (see Fig. \ref{fig:fitnessInteraction}), $b = 0.32, t = 9, p < 0.001$. In Expt.~1, the scripted label fit the hidden distractor just as well as the target, but in Expt.~2, the unconstrained labels fit the target much better and the hidden distractor much worse. In other words, the scripted labels used in Keysar's studies were less informative than speakers normally produce in this scenario.

Does differential informativity of scripted utterances account for the variability across items that we noted in Expt.~1? In Figure \ref{fig:itemLevel}, we compare the item-wise error \% and ratio of target fit to distractor fit. We find a correlation of $r = 0.76$ between these measures for Expt.~1, and a correlation of $r = 0.53$ for Expt.~2. The cross-experiment correlations of error rates from one experiment and label fitness ratios from the other, however, is only $r = 0.07$, suggesting that item-wise variability across the two experiments is driven by relative informativity of speaker utterances.

After separately establishing that matchers in Expt.~2 make fewer mistakes, that directors produce more informative utterances, and that informativity captures item-wise variability in error rates, we tested the link between these effects: does label informativity generally predict errors on critical trials? We used a mixed effects logistic regression model to estimate the effect of target and distractor fit on the probability of making a critical error in both experiments, including a random intercept for game ID. We found that participants are less likely to make errors when the target fit is higher, $b = -0.8, z = -4.1, p < 0.001$ and more likely to make errors when the distractor fit is higher, $b = 1.7, z = 3.7, p < 0.001$. Furthermore, a model including target fit and distractor fit in addition to item-level fixed effects is significantly better than a model including item alone, $\chi^2(2) = 36.2, p < 0.001$, implying that speaker informativity captures residual variance beyond the item-wise effects reported above.

\begin{figure}[t!]
\begin{center}
\includegraphics[scale=.5]{images/fitnessInteraction.pdf}
\caption{Mean fitness ratings provided by judges on Amazon Mechanical Turk. Error bars are bootstrapped 95\% confidence intervals.}
\label{fig:fitnessInteraction}
\end{center}
\end{figure}

@@

%consider the speaker's communicative goals and context. 
In our replication (Expt.~1), we found evidence that listeners neglect the speaker's epistemic state, as Keysar and colleagues claim.
We also found that when the speaker's utterance fit the hidden distractor better than the target, matchers were more likely to make errors. 
Indeed, we found that the extremely heterogeneous pattern of errors across items was well explained by the relative fit of the utterance to target and distractor.
This suggests that reference disambiguation was driven primarily by \emph{a priori} label fitness. 
%
However, this does not necessarily imply limitations on theory of mind---if speakers could be expected to naturally provide expressions which apply better to the target, then this behavior would be appropriate.
% If listeners arrive at this interpretation strategy through pragmatic reasoning, then neglecting the speaker's epistemic state would be rational and also the result of theory of mind use. 
In Expt.~2, we found that speakers did naturally produce more precise, informative utterances than required and these unconstrained utterances fit the target significantly better than they fit the hidden distractor. 
For example, no speaker in Expt.~2 produced ``the bottom block,'' which was used as a critical instruction in Expt.~1. Instead, they said ``the bluish block with a B'' or ``block with the blue writing,'' which relied on less confusable perceptual features and thereby decreased error rates. 
Thus, the errors observed in Expt.~1 can be explained as a result of an uncooperative confederate (speaker)---the experiment set up certain pragmatic expectations through the task context, then deliberately flouted them in critical trials. % Removing constraints on the speaker produced a dramatic reduction in overall matcher errors and systematic variation in error rates across items. We found that unscripted speakers systematically tended to produce more precise, costly referring expressions than their scripted counterparts -- for example, ``the clear audiocassette'' when there was only one ``tape'' in the speaker's display. These precise labels were less likely to fit the hidden distractor, and therefore less likely to lead to a matcher error.

@@

It is worth noting several significant differences between our study and \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
% that prevented a direct replication. 
The primary difference between our study and the original, of course, is that it was run on the web with participants connected through a virtual environment, instead of face-to-face in a room. We believe we addressed the major concern about exploring theory of mind in web experiments -- that participants do not truly believe they are interacting with another human -- by allowing instantaneous, responsive, real-time interaction.  On the other hand, it is known that textual communication, as in our chat box, can differ from face-to-face verbal communication, and there was no way of addressing this. % Directors watched objects move through the actual moment-by-moment mouse trajectories that matchers used to drag them, and matchers saw the actual timing of the messages sent by the director.

>>>>

\begin{abstract}
Theory of mind is a powerful cognitive ability: by the age of six, people are capable of accurately reasoning about others' beliefs and desires. An influential series of language understanding experiments by Keysar and colleagues, however, showed that adults systematically failed to take a speaker's beliefs into account, revealing limitations on theory of mind. In this paper we argue that these apparent failures are in fact successes. Through a minimal pair of replications comparing scripted vs. unscripted speakers, we show that critical utterances used by Keysar and colleagues are uncooperative: they are less informative than what a speaker would actually produce in that situation. When we allow participants to naturally interact, we find that listener expectations are justified and errors are reduced. This ironically shows that apparent failures of theory of mind are in fact attributable to sophisticated expectations about speaker behavior---that is, to theory of mind.

@@

\begin{figure*}[bt!]
\begin{center}
\includegraphics[scale = .37]{images/overall_view.png}
\end{center}
\vspace{-.25cm}
\caption{Interface used in the reported experiments. Objects behind the black squares were hidden from the director.}
\label{fig:interface}
\end{figure*}

@@

The argument offered by Keysar and colleagues is based on an elegant experimental paradigm, where participants played a simple communication game with a confederate. The two players were placed on opposite sides of a $4 \times 4$ grid containing a set of everyday objects (see Fig. \ref{fig:interface}). The confederate played the role of `director,' giving instructions about how to move objects around a grid, and the participant played the role of `matcher,' attempting to follow these instructions. For example, the objects in one trial included a cassette tape. The director gave an instruction like `move the tape up one square,' referring to the cassette.
Critically, some objects were occluded such that only the matcher could see them, creating an asymmetry in the players' knowledge. To perform accurately on critical trials, the matcher would need to apply theory of mind to reason about which objects were shared and which were private. 
For example, imagine a roll of tape were placed in an occluded slot: if a participant failed to account for the director's (partial) knowledge, she might interpret `tape' to mean the occluded roll of tape (which the director couldn't possibly know about). Indeed, \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} found that participants attempted to move the hidden item in 30\% of cases: 71\% of participants attempted to move this hidden item at least once (out of four critical cases) in the experimental condition, compared to 0\% in a control condition where there was no ambiguity over the referent. 
Additionally, eye-tracking data showed that participants considered the hidden item more often and for longer in the experimental condition than the control condition.

While these results are compelling, the paradigm has been criticized from few different angles. 
\citeA{HellerGrodnerTanenhaus08_Perspective} have pointed out that in many cases, the hidden object was a better fit for the referring expression than the one in common ground (e.g. the hidden roll of tape vs. the cassette tape for ``the tape''), making the hidden object \emph{a priori} more likely to be the referent; it would generally be fairer to compare two objects that fit the referring expression equally well. 
We validate this argument by empirically measuring relative fit of the expressions to the target and distractor items.
% Although relative fit was not empirically evaluated prior to our study,
Moreover, \citeA{HannaTanenhausTrueswell03_CommonGroundPerspective} argued that the viewpoint asymmetry paradigm is somewhat unnatural: common ground is typically built incrementally over the course of an interaction rather than presented all at once, and it is rare for a shared display to differ in perceptual accessibility. 
%Third, none of the experiments reported by Keysar and colleagues included a key comparison condition where the critical item (e.g. the roll of tape) was \emph{also} in common ground \cite{BrownSchmidtHanna11_IncrementalPerspectiveTaking}---perhaps people would refer to the critical item less in the privileged condition than in the full common ground condition, which would put a more positive spin on the results. %Note that none of these criticisms invalidate the paradoxical result that listeners move an occluded object, they just suggest an alternative explanation. Reference disambiguation is probabilistic: common ground and goodness of fit are weighed against one another.

In this paper, we offer an additional factor that helps account for Keysar's results. 
Theory of mind as applied within language understanding depends on an accurate model of what a speaker would say in different situations. Given an utterance, a listener can then reason backward to the most plausible situation \cite{Grice75_LogicConversation,Clark96_UsingLanguage,FrankGoodman12_PragmaticReasoningLanguageGames, GoodmanStuhlmuller13_KnowledgeImplicature}.
This suggests that we consider whether the utterances produced by the confederate in Keysar's critical conditions were actually what a speaker in that context would be expected to say. 
If not, then perhaps listeners are making choices that are in fact consistent with a correct pragmatic interpretation of the confederate's (uncooperative) utterance.
More precisely, when both players know that objects are occluded in the display, the speaker may tend to add additional precision to references in order to avoid confusion. If the listener \emph{expects} the speaker to do this, they will pragmatically pick the \emph{a priori} more likely referent of the referring expression, which in critical trials will be the occluded object. In other words, it is precisely \emph{because} the listener takes the speaker's mental state into consideration that they are tricked by an uncooperative confederate into choosing the wrong item. 
\begin{table*}
\begin{center}
\begin{tabular}{ p{2cm} | r | r |  r || r | r | r || r | r | r}
& \multicolumn{3}{c||}{\% attempted at least once} & \multicolumn{3}{c||}{\% attempted at least twice} & \multicolumn{3}{c}{\% of total cases}\\
\hline
& Orig. & Expt. 1 & Expt. 2 & Orig. & Expt. 1 & Expt. 2 & Orig. & Expt. 1 & Expt. 2  \\
Experimental & 71 & 93 & 61 & 46 & 57 & 32 & 30 & 43 & 24\\
Baseline        & 0   & 7   & 0   & 0   & 0   & 0   & 0   & 2   & 0\\
\end{tabular}
\caption{Side-by-side comparison of error rates in \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} and our two replications. The first two sections show the percentage of \emph{participants} attempting to move the occluded distractor at least once, or twice, of the four possible cases. The third section shows the percentage of \emph{all experimental trials} that the participant actually tried to move the occluded object. }
\label{table:mainResults}
\end{center}
\end{table*}

We began by replicating \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in a multi-player web experiment. We recruited participants to be both director and matcher (instead of using a confederate), but instructions for critical items, as well as a random subset of filler items, remained scripted as in the original study. We replicated the original finding, but noted a tendency of directors to be overinformative in unscripted filler trials. We then ran the same experiment without using any scripted instructions, observing unconstrained director utterances. We found much greater precision in unconstrained director utterances, which match targets much better than distractors, and better performance of the matchers. This minimal pair of experiments demonstrates that listener mistakes are at least partially due to the pragmatics of the task, ironically showing that apparent failures of theory of mind are in fact attributable to sophisticated expectations about speaker behavior---that is, to theory of mind.

@@

This paradigm differs from those used by \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in three primary ways. First, participants were not seated across from each other at a table: they each saw a view of the 4 x 4 grid on their screen and communicated via a text box. Second, we did not use a trained confederate. We randomly assigned one of the players to the role of the instruction giver, and maintained the original wording by scripting a subset of their instructions. Finally, the hidden object was not placed in a bag, in which respect our design more closely resembles \citeA{KeysarBarr___Brauner00_TakingPerspective}.

%The experiment proceeded through all eight items, with the director providing instructions and the participant moving the objects. Except for instructions with scripted messages, the participants were free to use the chat box for questions, small talk, or whatever else.
\begin{table*}
\begin{center}
\begin{tabular}{ m{1cm} | m{1.5cm} | m{1.5cm} | m{1.5cm} |  m{1.25cm} | m{1.6cm} | m{1.4cm} | m{1.1cm} | m{1.3cm} | m{1.3cm} |}
& & Item 1 & Item 2 & Item 3 & Item 4 & Item 5 & Item 6 & Item 7 & Item 8\\\hline
& instruction & ``glasses'' & ``bottom block'' & ``tape'' & ``large measuring cup'' & ``brush'' & ``eraser'' & ``small candle'' & ``mouse'' \\\hline
& target & sunglasses & block (3rd row) & cassette & medium cup & round hairbrush & board eraser & medium candle & computer mouse \\\hline
& hidden distractor & glasses case & block (4th row) & scotch-tape & large cup & flat \,\,\,hairbrush & pencil eraser & small candle & toy mouse \\
 \hline\hline
\multirow{2}{.75cm}{Expt.~1}  & \# incorrect & 0 & 5 & 1 & 3 & 2 &6 & 6 & 1 \\ 
                                               & \# correct & 6 & 3 & 1 & 8 & 2 & 2 & 4 & 6 \\
 \hline\hline
\multirow{2}{.75cm}{Expt.~2} & \# incorrect & 0 & 1 & 1 & 3 & 9 & 7 & 1 & 5 \\
                                              & \# correct   & 12&14&11&13&7 &8   &12& 8 \\

\end{tabular}
\vspace{-.5cm}
\caption{Item-wise error rates for critical trials}
\vspace{-.25cm}
\label{table:ItemWise}
\end{center}
\end{table*}

@@

As a proxy for the eye-tracking analyses reported by \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, we conducted a mouse-tracking analysis. We define the decision window as the span of time between the point when the matcher received their instruction message and when they started moving an object. If it took them multiple attempts to move the correct object, we restricted our analysis to the first attempt. We took the set of all mouse position samples in the decision window (collected approximately every 15ms), and measured what proportion were located inside the cell containing the target.  We had to exclude an additional 3 participants for this analysis, because the timestamps for director and matcher did not align and we could not establish the decision window properly.
We found that people tended to spend less total time hovering over the target cell on critical experimental trials than on baseline trials, $b = -.11, t(20) = -1.26, p = 0.224$, although perhaps due to an underpowered sample and imprecise measurements\footnote{This analysis includes participants who actually made errors, since the data is too sparse to exclude them.}, this result was not significant (see Fig. \ref{fig:hover}). %accounts for a significant proportion of variance $F(3, 52) = 10.7, p < 0.001$

By running this replication as a multi-player web experiment we have available an additional source of data beyond the original experiments: half of the instructions were unscripted, providing observations of natural production of referential descriptions for filler items. 
Informally, we noted a tendency toward additional, possibly unnecessary, precision in descriptions. Instead of ``move the stuffed animal down'', participants said ``move the stuffed panda bear down.'' Or, instead of saying ``move the plane to the right'' when there is only one plane, participants said ``move the red airplane to the right.'' 
Perhaps directors were taking the time to make more precise descriptions because they believed it was contextually relevant: both parties know that there are hidden objects in the environment increasing the chance of miscommunication from imprecise descriptions. 
If the matcher expected the director to be precise, then they would be justified in picking the first or best object that meets the description (rather than worrying excessively about occluded cells). 
That is, the scripted instructions used by the director for critical trials may have been \emph{uncooperative} for this situation, and thus led matchers astray. 
We tested this prediction in Expt.~2, where we removed the scripted instructions and allowed speakers to refer to items however they wished.
By the reasoning above we expected to see more precise descriptions by unscripted directors and fewer errors by matchers in the critical trials.

@@

 \begin{figure}[b!]
\begin{center}
\includegraphics[scale=.5]{images/mousetracking.pdf}
\vspace{-.5cm}

\caption{Mean percentage of time spent hovering over target cell in the experimental and baseline conditions for Expt.~1 (left) and Expt.~2 (right). Error bars are 95\% confidence intervals.}
\label{fig:hover}
\end{center}
\end{figure}

@@

We find that patterns of errors in Expt.~2 diverge significantly from a uniform distribution across items, $\chi^2(7) = 24.8, p < 0.001$. Looking more closely at these patterns, we see that the only items where errors are consistently made are those where the more precise utterances used remain ambiguous. For example, in item 5, both the target and distractor are hair brushes: one is round and one is flat. Many participants produced the sub-class label ``hair brush,'' which was more precise than the scripted basic-level ``brush'' from Expt.~1, but the two objects were still confusable at the sub-class level. If we remove the two most difficult items (the ``hair brushes'' and the ``erasers''), the total percentage of errors on experimental trials drops from 24\% to 10\% and the percentage of participants making at least one error drops from 61\% to 32\%.

When we conducted a mouse-tracking analysis identical to the one reported for Expt.~1, we again found a decrease in total hover time over the target in experimental vs. baseline trials $b = 0.07, t(54) = -2.1, p = 0.045$, indicating that the presence of a hidden distractor still interfered with participants directly choosing the target. While the effect appears to be slightly attenuated compared to Expt.~1 (see Figure \ref{fig:hover}), there is no significant interaction of experiment and condition on hover time $b = 0.04, t(74) = 0.45, p = 0.657$. Again, given the low sample size in Expt.~1, the noise from error, and the noise from counting samples instead of differences in timestamps, this result is limited.
\begin{figure}[t!]
\begin{center}
\includegraphics[scale=.5]{images/itemwisefitness.pdf}
\vspace{-.25cm}
\caption{Item-wise error rates compared with label fitness ratios, for Expt.~1. Fitness ratios are scaled to the unit interval in order to visually compare to error rates. Error bars are bootstrapped 95\% confidence intervals for fitness and 95\% highest density intervals for error rates.}
\vspace{-.25cm}
\label{fig:itemLevel}
\end{center}
\end{figure}

Next, we test whether these improvements in performance are in fact due to more informative speaker behavior. We recruited twenty judges on Amazon Mechanical Turk, who provided ratings for how well the 71 unique labels used by speakers across both experiments (including scripted labels) fit the target and hidden distractor objects. Their responses were given on a slider with endpoints labeled ``not at all'' and ``perfectly.''  Inter-rater reliability was relatively high, with intra-class correlation coefficient of $0.6$ (95\% CI = $[0.54, 0.66]$). In a mixed model including random intercepts for raters and items, we found a significant crossover interaction of experiment and referent on mean fitness rating (see Fig. \ref{fig:fitnessInteraction}), $b = 0.32, t = 9, p < 0.001$. In Expt.~1, the scripted label fit the hidden distractor just as well or better than the target, but in Expt.~2, the unconstrained labels fit the target much better and the hidden distractor much worse. In other words, the scripted labels used in Keysar's studies were less informative than speakers normally produce in this scenario.

Does differential informativity of scripted utterances account for the variability across items that we noted in Expt.~1? In Figure \ref{fig:itemLevel}, we compare the item-wise error \% and ratio of target fit to distractor fit. We find a correlation of $r = 0.76$ between these two measures for Expt.~1, and a correlation of $r = 0.53$ for Expt.~2. Comparing error rates from one experiment with label fitness ratios from the other, however, yields much weaker correlations ($r = 0.07$), suggesting that item-wise variability across the two experiments is driven by relative informativity of speaker utterances.

After separately establishing that matchers in Expt.~2 make fewer mistakes, that directors in Expt.~2 produce more informative utterances, and that informativity captures item-wise variability in error rates in both experiments, we tested the link between these effects in our aggregated data: does label informativity generally predict errors on critical trials? We used a mixed effects logistic regression model to estimate the effect of target and distractor fit on the probability of making a critical error in both experiments, including a random intercept for game ID. We found that participants are less likely to make errors when the target fit is higher, $b = -0.8, z = -4.1, p < 0.001$ and more likely to make errors when the distractor fit is higher, $b = 1.7, z = 3.7, p < 0.001$. Furthermore, a model including target fit and distractor fit in addition to item-level fixed effects is significantly better than a model including item alone, $\chi^2(2) = 36.2, p < 0.001$, implying that speaker informativity captures residual variance beyond the item-wise effects reported above.

\begin{figure}[t!]
\begin{center}
\includegraphics[scale=.5]{images/fitnessInteraction.pdf}
\vspace{-.25cm}
\caption{Mean fitness ratings provided by judges on Amazon Mechanical Turk. Error bars are bootstrapped 95\% confidence intervals.}
\vspace{-.25cm}
\label{fig:fitnessInteraction}
\end{center}
\end{figure}

@@

%consider the speaker's communicative goals and context. 
In our replication (Expt.~1), we found evidence that listeners neglect the speaker's epistemic state, as Keysar and colleagues claim.
We also found that when the speaker's utterance fit the hidden distractor better than the target, matchers were more likely to make errors. 
Indeed, we found that the extremely heterogeneous pattern of errors across items was well explained by the relative fit of the utterance to target and distractor.
This suggests that reference disambiguation was driven primarily by \emph{a priori} label fitness rather than consideration of occluded vs. mutual visible items. 
%
However, this does not necessarily imply limitations on theory of mind---if speakers could be expected to naturally provide expressions which apply better to the target, then this behavior would be appropriate.
% If listeners arrive at this interpretation strategy through pragmatic reasoning, then neglecting the speaker's epistemic state would be rational and also the result of theory of mind use. 
In Expt.~2, we found that speakers did naturally produce more precise, informative utterances than required and these unconstrained utterances fit the target significantly better than they fit the hidden distractor. 
For example, no speaker in Expt.~2 produced ``the bottom block,'' which was used as a critical instruction in Expt.~1. Instead, they said ``the bluish block with a B'' or ``block with the blue writing,'' which relied on less confusable perceptual features and thereby decreased error rates. 
Thus, the errors observed in Expt.~1 can be explained as a result of an uncooperative confederate (speaker)---the experiment set up certain pragmatic expectations through the task context, then deliberately flouted them in critical trials. % Removing constraints on the speaker produced a dramatic reduction in overall matcher errors and systematic variation in error rates across items. We found that unscripted speakers systematically tended to produce more precise, costly referring expressions than their scripted counterparts -- for example, ``the clear audiocassette'' when there was only one ``tape'' in the speaker's display. These precise labels were less likely to fit the hidden distractor, and therefore less likely to lead to a matcher error.

@@

It is worth noting several significant differences between our study and \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
The primary difference between our study and the original, of course, is that it was run on the web with participants connected through a virtual environment, instead of face-to-face in a room. We believe we addressed the major concern about exploring theory of mind in web experiments -- that participants do not truly believe they are interacting with another human -- by allowing instantaneous, responsive, real-time interaction.  On the other hand, it is known that textual communication, as in our chat box, can differ from face-to-face verbal communication, and there was no way of addressing this. % Directors watched objects move through the actual moment-by-moment mouse trajectories that matchers used to drag them, and matchers saw the actual timing of the messages sent by the director.
